---
title: "Downloader for National Minimum dataset for Social Care Data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Intro

Short Rmd to take care of downloading and converting the Skills for Care National Minimum Data Set.

Skills for Care's National Minimum Data Set for Social Care (NMDS-SC) is recognised as the leading source of robust workforce intelligence for adult social care. The NMDS-SC collects information online about providers offering a social care service and their employees. Social care providers can register, maintain and access their business information at www.nmds-sc-online.org.uk. 

https://www.nmds-sc-online.org.uk/Get.aspx?id=/Research/Adult%20social%20care%20workforce%20reports/Open%20data/Open%20Data%20August%202017.csv

Define URL and dates for each of the csv's
```{r}
root <- 'https://www.nmds-sc-online.org.uk/Get.aspx?id='

dates <- c(
  'Aug 17',
  'Feb 17',
  'Aug 16',
  'Feb 16'
)

files <- c(
'/Research/Adult social care workforce reports/Open data/Open Data August 2017.csv',
'/Reports/Accessing data/NMDSSC Open Data February 2017.csv',
'/Research/Adult social care workforce reports/Open data/NMDS-SC Open Data Analysis File - August 2016.csv',
'/Research/Adult social care workforce reports/Open data/NMDS-SC Open Data Analysis File - February 2016.csv')

```

# Download  files 
(Note mode=wb to correctly download binary files.)
```{r}
srcdat="../../Source Data/"

for(i in 1:length(files)){
  
  download.file(url=paste0(root,files[i]), paste0(srcdat,dates[i],'.csv'), mode="wb")
}
```

#Read in Downloaded Files
```{r}
Aug17<-read.csv(paste0(srcdat,dates[1],'.csv'))
```

Ok looks like there is lots of data in here! Need to digest the code book:
<https://www.nmds-sc-online.org.uk/Get.aspx?id=834314>

... ok unfortunately looks like the only geographic field is 'Region' - No LA level data here (though it must exist because it is in the dashboard on the website). There is a lot of information to be had though, definite potential for less geographically based analyses. Should also investigate if the LA level data is on the UK Data Store website.