---
title: "Tall To Wide for LocalArea Data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("~/Dropbox/Current Work/BRC IMD/Data/LocalHealth")
```

Data downloaded from the localhealth.org.uk website comes in the tall tidy data format (a column dictating the variable, with others dictating the value and confidence interval.)
For the purpose of using it in a database style with joins in QGIS it is more convenient to have it in a wide format

This script will do this.

The first step is to read the data all into R
```{r Read in CSV's}

T1data=read.csv('Theme1_2017.csv', header=TRUE)
T2data=read.csv('Theme2_2017.csv', header=TRUE)
T3data=read.csv('Theme3_2017.csv', header=TRUE)
T4data=read.csv('Theme4_2017.csv', header=TRUE)
```

Next we load up dplyr to manipuate these data tables
```{r}
library(dplyr)
library(tidyr)
```

We want to trim the tables down a bit, they include 
```{r}

T1data <- T1data %>%
  mutate(id_Name=paste(as.character(ThemeID),Name)) %>%
  select(id_Name, AreaCode, Value) 

T2data <- T2data %>%
  mutate(id_Name=paste(as.character(ThemeID),Name)) %>%
  select(id_Name, AreaCode, Value) 

T3data <- T3data%>%
  mutate(id_Name=paste(as.character(ThemeID),Name)) %>%
  select(id_Name, AreaCode, Value) 

T4data <- T4data%>%
  mutate(id_Name=paste(as.character(ThemeID),Name)) %>%
  select(id_Name, AreaCode, Value) 

```


At this point its probably a good idea to group all the data into one table to avoid unneceary repetition. - Note that the ThemeID column is kept
```{r Combine into a single table}
T_data <- rbind(T1data, T2data, T3data, T4data)

```


Next, we want to reorganise this from a tall format to a wide format. This can be done using the tidyr function spread.
```{r}
Wide_T_data <- T_data %>%
  spread(id_Name,Value)
```

Now that we have our wide data we want to write it back out to a CSV for use in Q
```{r}
filename='WideLocalHealthData.csv'
write.csv(Wide_T_data,filename, append=FALSE)
```

