```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load data handling libraries
At this point I will also load a few other useful libraries:
```{r}
library(dplyr)
library(jsonlite)
```

# Search - does not work??
https://www.nomisweb.co.uk/api/v01/dataset/def.sdmx.xml?search=contenttype-<contenttype>  #- Format
https://www.nomisweb.co.uk/api/v01/dataset/def.sdmx.xml?search=contenttype-sources-jsa    #- Gets all the JSA datasets
https://www.nomisweb.co.uk/api/v01/dataset/def.htm?search=contenttype-sources-jsa         #- Gets all the JSA datasets

# def.sdmx.xml

# Returns a description but not an index of the content type
https://www.nomisweb.co.uk/api/v01/contenttype/sources/id/jsa.xml #- Job Seekers Allowance
https://www.nomisweb.co.uk/api/v01/contenttype/sources/id/cc.xml  #- Claimant Count

# Gets the contenttype in a json for a given dataset using the Overview endpoint. This is not standard sdmx.
https://www.nomisweb.co.uk/api/v01/dataset/nm_1_1.overview.json?select=contenttypes

# Get a JSON containing all contenttype id's which fall under the AnnotationTitle 'contenttype/sources'
```{r}
contenttypejs<-"https://www.nomisweb.co.uk/api/v01/contenttype/sources.json"

json_data <- fromJSON(file=contenttypejs)
```





# Returns a list (of 1) of all the ways there are of grouping data. This would appear to just be 'sources'
/api/v01/contenttype/index.xml



