---
title: "Support at Home Data"
author: "Matthew Malcher"
date: "13/12/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Prerequisites.
This file requires that all chunks in `SDMX Access.Rmd` be run first to define all the functions and load all the relevant libraries for accessing data via the Nomis API.

we also want to load the following additional libraries:
```{r}
library(tidyr)#for spreading datasets from tall to wide
library(xlsx) #for writing out to spreadsheet
```

# Support at Home
The aim of this project is to curate a selection of geographical UK datasets relevant to NGO's. To help focus on how to do this, and what might actually be useful it is probably sensible to do a few case studies.

## Find JSA variables


```{r}
available_OAs(ap,qformat,'NM_1_1','2092957697')%>% filter(id =='2092957697TYPE464')

 available_OAs(ap,qformat,'NM_7_1','2092957697')%>% filter(id =='2092957697TYPE464')

 available_OAs(ap,qformat,'NM_10_1','2092957697') %>% filter(id =='2092957697TYPE464')
```

We can see that all three variables are available at local authority level. Good!

## Get Output Areas
Now we can get the list of output areas available for each dataset.
```{r}
OAlist<- list_OAs(ap,qformat,'NM_1_1','2092957697','TYPE464')
```

# NM_1_1 - JSA Claimants

##  Get Structure
```{r}
get_struct (ap,qformat,'NM_1_1')
```





##  Get Items
```{r}
get_items(ap,'NM_1_1',qformat)
```
##  Get Measures
```{r}
get_measures(ap,qformat,'NM_1_1','2092957697','TYPE464')
```

##  Get Data
```{r}
qlist<-list(geography='2092957697TYPE464',      # Local Authority Areas
            sex=7,                              # Genders: Total for both Genders
            item=1,                             # Item: Total Claimints
            measures='20100,20201,20202,20203', # Measures: claimants, workforce,
            time='latest')                      # Latest Time

data_NM_1_1<-get_data_sdmx(qlist, 'NM_1_1')

```

## Tidy Data
We notice that we get 1624 rows for NM_1_1:

  * 20100 People Claming,
  * 20201 workplace estimate, 
  * 20202 Economically Active 
  * 20203 People Claiming as a proportion of resident population

4 measures x 406 regions = 1624 rows

Our target is a wide dataset we can join in QGIS or similar with 1 row per geography. Therefore we want to use something like spread from tidyr to tidy this up. We can do this using the measures column in this case.
```{r}
#  Spread Data Frame
data_NM_1_1 <- data_NM_1_1 %>%
  select(GEOGRAPHY, MEASURES, OBS_VALUE) %>%
  spread(key='MEASURES',  value='OBS_VALUE' ) 

# Rename Spread Columns
data_NM_1_1 <- reaname_via_codelist(data_NM_1_1,'CL_1_1_MEASURES' )
```


# NM_7_1 - claimant count - occupation age and duration

```{r}
info_on_dataset(ap,qformat,'NM_7_1')
```

##  Get Structure
```{r}
get_struct (ap,qformat,'NM_7_1')
```

## Code Lists for non standard quantities
```{r}
get_codelist(cl,qformat,'CL_7_1_OCCUPATION')
get_codelist(cl,qformat,'CL_7_1_AGE_DUR')
```
Note that there are 112 possible occupations and 238 age_duration bands. As such, if you dont restrict one or both of these, with 406 output areas requested and data returned in a tall format you are gonna get pretty complex and hit the 25,000 cell limit pretty quickly.

As another way of thinking about it, you could have 60 data items for each of the 406 Local Authorities without going over the limit. (Though I am unsure if the additional default columns which come through count towards the limit.)

## Get Items
```{r}
get_items(ap,'NM_7_1',qformat)
```

## Get Measures
```{r}
get_measures(ap,qformat,'NM_7_1','2092957697','TYPE464')
```


## Get Data
Get NM_7_1 Data - This is probably the most basic request you could make, with both age_duration and occupation set to their 'total' options. There is a lot of available complexity and subsetting you can do with this dataset!
```{r}
qlist<-list(geography='2092957697TYPE464',              # Local Authority Areas
            sex=7,                                      # Genders: Total for both Genders
            item=1,                                     # Item: Sought Occupation
            measures='20100',                           # Measures: value
            time='latest',                              # Latest Time
            AGE_DUR='1,2,3,4,5,6,7,8,9,10,11,12,13',    # Select the different age bands
            Occupation = 589824)                        # All Occupations                

data_NM_7_1 <-get_data_sdmx(qlist, 'NM_7_1')
```

## Tidy Data 
```{r}
# Spread data frame by age bands
data_NM_7_1 <- data_NM_7_1 %>%
  select(GEOGRAPHY, AGE_DUR, OBS_VALUE) %>%
  spread(key='AGE_DUR',  value='OBS_VALUE' ) 

# Rename using codelist for age bands
data_NM_7_1<-reaname_via_codelist(data_NM_7_1,'CL_7_1_AGE_DUR' )
```

# NM_10_1 - Jobseeker's Allowance flows

##  Get Structure
```{r}
get_struct (ap,qformat,'NM_10_1')
```
## Code Lists for non standard quantities
```{r}
get_codelist(cl,qformat,'CL_10_1_FLOW')
```


## Get Items
```{r}
get_items(ap,'NM_10_1',qformat)
```

## Get Measures
```{r}
get_measures(ap,qformat,'NM_10_1','2092957697','TYPE464')
```

## Get Data
Get NM_10_1 Data
```{r}
qlist<-list(geography='2092957697TYPE464',              # Local Authority Areas
            sex=7,                                      # Genders: Total for both Genders
            item=1,                                     # Item:  Total Claimints
            measures='20100',                           # Measures: value
            time='latest',                              # Latest Time
            FLOW='1,2,3')                               # Get Stock, In & Out Flows

data_NM_10_1 <-get_data_sdmx(qlist, 'NM_10_1')
```


## Tidy Data 

```{r}
# Spread data frame by age bands
data_NM_10_1 <- data_NM_10_1 %>%
  select(GEOGRAPHY, FLOW, OBS_VALUE) %>%
  spread(key='FLOW',  value='OBS_VALUE' ) 

# Rename using codelist for age bands
data_NM_10_1<-reaname_via_codelist(data_NM_10_1,'CL_10_1_FLOW' )
```

# Export Data

## Combine Data Frames
```{r}
jsa_data <- inner_join(data_NM_1_1, data_NM_7_1, by = "Geography")
jsa_data <- inner_join(jsa_data, data_NM_10_1, by = "Geography")
```

## Add Geography Names
We can join our Geography names from the output of list_OAs to our data obtained with get_data
```{r}
OAlist_NM_1_1 <- list_OAs(ap,qformat,'NM_1_1','2092957697','TYPE464')

jsa_data <-left_join(OAlist_NM_1_1,jsa_data, by = "GEOGRAPHY")
```

## Write out to csv / excel
Write out to csv
```{r}
filename='../../Processed Data/WideNomiswebJsaData.csv'
write.csv(jsa_data,filename, append=FALSE)
```

Write out to Excel
```{r}
# Write the weights and total weight to one worksheet
# write.xlsx2(NomisDatasets, 'NomisDatasets.xlsx', sheetName="Datasets", 
#   col.names=TRUE, row.names=FALSE, append=FALSE, showNA=TRUE)
```
