---
title: "SDMX"
author: "Matthew Malcher"
date: "17/11/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load data handling libraries
At this point I will also load a few other useful libraries:
```{r}
library(dplyr)
```

#  Load rsdmx
There is an R library for working with SDMX. NOMIS is included as a predefined data source! This looks like it might make things a bit easier, but I cant figure out how to take advantage of it. Fortunately you can also buld your own URL's and feed them to the `readSDMX` command provided by the library.
```{r RSDMX}
library(rsdmx)
as.data.frame(getSDMXServiceProviders()) %>% filter(agencyId=='NOMIS')
```

The script seems to have some issues when run on Windows 10 - I susepect this is to do with http/https and ssl RC4 being depracated. The following is an attempt to mess around with the RCurl settings to ignore certificate errors. I dont know if it works 
```{r Rcurl with settings for https on windows}
library(RCurl)
# curlprefs <- curlOptions(ssl.verifypeer=FALSE, ssl.verifyhost=FALSE)
# curlSetOpt(.opts=curlprefs, curl = getCurlHandle())
```

There are some common bits of info used in all requests 
```{r Set API ap and Formats}
# Define API access point
ap<-'https://www.nomisweb.co.uk/api/v01/dataset' 

# Define the format of the data returned when interogating Nomis. The rsdmx package allows SMDX in XML format to be read into data frames.
qformat <-'def.sdmx.xml'

# Define the format of the data returned when downloading Files. The rsdmx package allows SMDX in XML format to be read into data frames.
dformat<- 'compact.sdmx.xml' # For some reason if I use generic.sdmx.xml I get back partial XML files?
# using compact sdmx gives a URI error but seems to return the correct data

```


# Define Functions for Interacting with Nomis API

```{r Define Functions}
get_all_dataset<-function(ap,qformat){
    queryUrl <- paste(ap,qformat,sep='/')
    all_data <- as.data.frame(readSDMX(queryUrl))
    return(all_data)
  }

info_on_dataset <- function(ap,qformat,dataset){
  queryUrl<-paste0(paste(ap,dataset,sep='/'),'.',qformat)
  # print('Querying: ',queryUrl)
  SingleDataset <- as.data.frame(readSDMX(queryUrl))
  return(SingleDataset)
 }

available_regions <- function(ap,qformat,dataset){
  qtype<-'geography'
  queryUrl<-paste0(paste(ap,dataset,qtype,sep='/'),'.',qformat)
  regionlist <- as.data.frame(readSDMX(queryUrl))
  return(regionlist)
}

get_genders <-function(ap,qformat,dataset){
  qtype<-'sex'
  queryUrl<-paste0(paste(ap,dataset,qtype,sep='/'),'.',qformat)
  genders <- as.data.frame(readSDMX(queryUrl))
  return(genders)
}

available_OAs <- function(ap,qformat,dataset,regioncode){
  qtype<-'geography'
  queryUrl<-paste0(paste(ap,dataset,qtype,regioncode,sep='/'),'.',qformat)
  availableOAlist <- as.data.frame(readSDMX(queryUrl))
  return(availableOAlist)
}

list_OAs <- function(ap,qformat,dataset,regioncode,OAtype){
  qtype<-'geography'
  queryUrl<-paste0(paste(ap,dataset,qtype,regioncode,sep='/'),OAtype,'.',qformat)
  OAlist <- as.data.frame(readSDMX(queryUrl))
  OAlist<-rename(OAlist, GEOGRAPHY=id) # Rename the id column to GEOGRAPHY to match output of 
  return(OAlist)
}

get_items <- function(ap,dataset,qformat){
  
  qtype<-'item'
  queryUrl<-paste0(paste(ap,dataset,qtype,sep='/'),'.',qformat)
  AvailableItems <- as.data.frame(readSDMX(queryUrl))
  return(AvailableItems)
}

get_times <- function(ap,dataset,qformat){
  
  qtype<-'time'
  queryUrl<-paste0(paste(ap,dataset,qtype,sep='/'),'.',qformat)
  AvailableTimes <- as.data.frame(readSDMX(queryUrl))
  return(AvailableTimes)
}

get_measures <- function(ap,qformat,dataset,regioncode,OAtype){
  
  qtype<-'measures'
  queryUrl<-paste0(paste(ap,dataset,qtype,regioncode,sep='/'),OAtype,'.',qformat)
  print(queryUrl)
  measures <- as.data.frame(readSDMX(queryUrl))
  return(measures)
}


get_data <- function(dataset, geography, sex, item, measures, dformat , time){
  
  # Pre enter the valid Time options
  timeopts=c(
    "latest", # the latest available data for this dataset
    "previous", # the date prior to "latest"
    "prevyear", # the date one year prior to "latest"
    "first") # the oldest available data for this dataset
  
  # Pre enter the valid Time options
  sexopts=c("5",  #Male Only
            "6",  #Female Only
            "7",  #Total
            "5,6" #Both
            )
  
  query<-paste0('?geography=',geography,'&sex=',sexopts[sex],'&item=',item,'&measures=',measures,'&time=', timeopts[time])
  
  queryUrl<-paste0(paste(ap,dataset,sep='/'),'.',dformat,query)
  print(queryUrl)
  df <- as.data.frame(readSDMX(queryUrl))
  
  return(df)
}
```

Now, lets get started with the nomisweb API!

With the following commands we can get a data frame with details of all the datasets available on NOMIS
```{r}
NomisDatasets <- get_all_dataset(ap,qformat)
```


Alternatively, if we know which dataset we want then we can request just that
```{r}
info_NM_1_1 <- info_on_dataset(ap,qformat,'NM_1_1')
info_NM_1_2 <- info_on_dataset(ap,qformat,'NM_1_2')

```


Get a Table of available regions for a dataset:
```{r}
regions_NM_1_1<-available_regions(ap,qformat,'NM_1_1')
```

Get a Table of available output areas for a dataset within a region identified by its id
```{r}
OAs_NM_1_1 <- available_OAs(ap,qformat,'NM_1_1','2092957697')
```

https://www.nomisweb.co.uk/forum/posts.aspx?tID=555&fID=2

Get List of output areas for a given dataset, region and output area type
```{r}
OAlist_NM_1_1 <- list_OAs(ap,qformat,'NM_1_1','2092957697','TYPE464')

```

Get a Table of available sex information for a dataset
```{r}
dataset<-'NM_1_1'
genders_NM_1_1 <- get_genders(ap,qformat,dataset)

```

Get the measures available for a given geography - these are denominators - i.e. out of the number of claimints, the number of people living there etc.
```{r}
measures_NM_1_1 <- get_measures(ap,qformat,'NM_1_1','2092957697','TYPE464')
```

Get available 'items' for a specified dataset (i.e. the variables)
```{r}
dataset<-'NM_1_1'
items_NM_1_1 <- get_items(ap,dataset,qformat)
```

Get available 'times' for a specified dataset 
```{r}
dataset<-'NM_1_1'
times_NM_1_1 <- get_times(ap,dataset,qformat)

```


Get available data for a specified dataset
```{r}
dataset<-'NM_1_1'
geography='2092957697TYPE464'
geography='2092957697'
sex=3
item='1'
time=1
measures='20100,20201,20202,20203'
measures='20100'

data_NM_1_1 <- get_data(dataset, geography, sex, item, measures, dformat, time)

```

We can join our Geography names from the output of list_OAs to our data obtained with get_data
```{r}
left_join(filter(df, SEX=='7'), OAlist_NM_1_1, by = "GEOGRAPHY")
```

