---
title: "SDMX"
author: "Matthew Malcher"
date: "17/11/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load data handling libraries
At this point I will also load a few other useful libraries:
```{r}
library(dplyr)
# library(tidyr)
```


#  Load rsdmx
There is an R library for working with SDMX. NOMIS is included as a predefined data source! This looks like it might make things a bit easier, but I cant figure out how to take advantage of it. Fortunately you can also buld your own URL's and feed them to the `readSDMX` command provided by the library.
```{r }
library(rsdmx)
providers <- getSDMXServiceProviders();
as.data.frame(providers) %>% filter(agencyId=='NOMIS')
```

The script seems to have some issues when run on Windows 10 - I susepect this is to do with http/https and ssl RC4 being depracated. The following is an attempt to mess around with the RCurl settings to ignore certificate errors. I dont know if it works 
```{r Rcurl with settings for https on windows}
library(RCurl)
curlprefs <- curlOptions(ssl.verifypeer=FALSE, ssl.verifyhost=FALSE)
curlSetOpt(.opts=curlprefs, curl = getCurlHandle())
```


There are some common bits of info used in all requests 
```{r}
# Define API access point
ap<-'https://www.nomisweb.co.uk/api/v01/dataset' 

# Define the format of the data returned. The rsdmx package allows SMDX in XML format to be read into data frames.
qformat <-'def.sdmx.xml'

```

Now, lets get started with the nomisweb API!

With the following commands we can get a data frame with details of all the datasets available on NOMIS
```{r}
get_all_dataset<-function(ap,qformat){
    queryUrl <- paste(ap,qformat,sep='/')
    all_data <- as.data.frame(readSDMX(queryUrl))
    return(all_data)
  }

NomisDatasets <- get_all_dataset(ap,qformat)
```


Alternatively, if we know which dataset we want then we can request just that
```{r}

info_on_dataset <- function(ap,qformat,dataset){
  queryUrl<-paste0(paste(ap,dataset,sep='/'),'.',qformat)
  # print('Querying: ',queryUrl)
  SingleDataset <- as.data.frame(readSDMX(queryUrl))
  return(SingleDataset)
}

info_NM_1_1 <- info_on_dataset(ap,qformat,'NM_1_1')
info_NM_1_2 <- info_on_dataset(ap,qformat,'NM_1_2')

```


Get a Table of available regions for a dataset:
```{r}

available_regions <- function(ap,qformat,dataset){
  qtype<-'geography'
  queryUrl<-paste0(paste(ap,dataset,qtype,sep='/'),'.',qformat)
  regionlist <- as.data.frame(readSDMX(queryUrl))
  return(regionlist)
}

regions_NM_1_1<-available_regions(ap,qformat,'NM_1_1')

```

Get a Table of available output areas for a dataset within a region identified by its id
```{r}
available_OAs <- function(ap,qformat,dataset,regioncode){
  qtype<-'geography'
  queryUrl<-paste0(paste(ap,dataset,qtype,regioncode,sep='/'),'.',qformat)
  availableOAlist <- as.data.frame(readSDMX(queryUrl))
  return(availableOAlist)
}

OAs_NM_1_1 <- available_OAs(ap,qformat,'NM_1_1','2092957697')
```

https://www.nomisweb.co.uk/forum/posts.aspx?tID=555&fID=2

Get List of output areas for a given dataset, region and output area type
```{r}

list_OAs <- function(ap,qformat,dataset,regioncode,OAtype){
  qtype<-'geography'
  queryUrl<-paste0(paste(ap,dataset,qtype,regioncode,sep='/'),OAtype,'.',qformat)
  OAlist <- as.data.frame(readSDMX(queryUrl))
  return(OAlist)
}

OAlist_NM_1_1 <- list_OAs(ap,qformat,'NM_1_1','2092957697','TYPE464')

```

Get a Table of available sex information for a dataset
```{r}
get_genders <-function(ap,qformat,dataset){
  qtype<-'sex'
  queryUrl<-paste0(paste(ap,dataset,qtype,sep='/'),'.',qformat)
  genders <- as.data.frame(readSDMX(queryUrl))
  return(genders)
}

dataset<-'NM_1_1'
genders_NM_1_1 <- get_genders()

```

Get the measures available for a given geography - these are denominators - i.e. out of the number of claimints, the number of people living there etc.
```{r}
get_measures <- function(ap,qformat,dataset,regioncode,OAtype){
  
  qtype<-'measures'
  queryUrl<-paste0(paste(ap,dataset,qtype,regioncode,sep='/'),OAtype,'.',qformat)
  measures <- as.data.frame(readSDMX(queryUrl))
  return(measures)
}


measures_NM_1_1 <- get_measures(ap,qformat,'NM_1_1','2092957697','TYPE464')
```

Get available 'items' for a specified dataset (i.e. the variables)
```{r}
dataset<-'NM_1_1'
qtype<-'item'
queryUrl<-paste0(paste(ap,dataset,qtype,sep='/'),'.',qformat)
queryUrl
AvailableItems <- as.data.frame(readSDMX(queryUrl))
AvailableItems

```

/api/v01/dataset/NM_1_1.generic.sdmx.xml?geography=2038432081&sex=5&item=1&measures=20100&time=latest

Get available 'items' for a specified dataset (i.e. the variables)
```{r}

dataset<-'NM_1_1'

geography='2092957697TYPE464'
sex='7'
item='1'
measures='20100'
time='latest'

query<-paste0('?geography=',geography,'&sex=',sex,'&item=',item,'&measures=',measures,'&time=latest')

#Dont need to have all items  
query<-paste0('?geography=',geography,'&item=',item,'&measures=',measures,'&time=latest')

format<-'.generic.sdmx.xml'

file<-paste0(paste(ap,dataset,sep='/'),format,query)

sdmx <- readSDMX(file)

df <- as.data.frame(sdmx)

df
```

```{r}

df %>%
  filter(SEX=='7') %>%
  join() %>%
  
```

