---
title: "MSOA to LAD"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load up required libraries.
```{r}
library(dplyr)
library(tidyr)
```

In a previous script we downloaded the Theme summary csv's and reorganised them from a tall to a wide format. The resulting csv, as with the source local health data is provided at the MSOA level.

This script will take that MSOA data and group it by Local Authority Region

https://ons.maps.arcgis.com/home/item.html?id=ef72efd6adf64b11a2228f7b3e95deea

The first step is to read MSOA Local Health data into R:
```{r Read in CSV's}
LocalHealthMSOAdata=read.csv('../Processed Data/WideLocalHealthData.csv', header=TRUE)
```

Next we need a way to translate from MSOA's to LAD's. Fortunatley, at their Open Geography Portal (http://geoportal.statistics.gov.uk) the ONS has provided some lookup tables.

Here we will use the Postcode <> LSOA <> MSOA <> LAD (2011) lookup table. This can be found at: https://ons.maps.arcgis.com/home/item.html?id=ef72efd6adf64b11a2228f7b3e95deea

To provide some resilience, a zip file containing this table (PCD11_OA11_LSOA11_MSOA11_LAD11_EW_LU_aligned_v2.zip) is kept in this repo.

Lets read it in.
```{r}
# Use unz to read in the file from inside the zip - otherwise quite large
LookupTable <- read.csv(unz(description = '../Source Data/PCD11_OA11_LSOA11_MSOA11_LAD11_EW_LU_aligned_v2.zip',filename="PCD11_OA11_LSOA11_MSOA11_LAD11_EW_LU_aligned_v2.csv"))

```

Ok, thats a big list - 1.3m postcodes. We only care about LAD's and MSOA's so we can cut out a few columns, and remove the duplicate rows:
```{r}
# reduce table to only relevant columns
LookupTable<-LookupTable %>% select(MSOA11CD,MSOA11NM,LAD11CD,LAD11NM)

# Get rid of duplicate rows
LookupTable<-unique(LookupTable)  

filename='../Processed Data/MSOA2LAD.csv'
write.csv(LookupTable,filename, append=FALSE)
```

At this point it would be good to check that the lookup table is right. We can do this is to get MSOA shapefile, join it to the lookup table using the MSOA codes, then colour the MSOA areas by LAD and overlay against the LAD layer to visually check. We can see that everything looks [good!](https://github.com/mattmalcher/IndexOfNeed/blob/master/Datasets/LocalHealth/Maps/MSOAs_in_LADs.png)
