---
title: "Nomisweb API Access"
author: "Matthew Malcher"
date: "14/11/2017"
output: html_document
---

# Setup

##Introduction
This markdown document contains multiple code snippets for interacting with the Nomisweb API using R. The idea is that these examples are set up so that they can readily be repurposed.

## Load Dependencies
Set up knitr:
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Load the libraries we will be using:
```{r load_libs_1, echo=FALSE}
library(RCurl)
library(dplyr)
library(jsonlite)
```


## Documentation and References
The documentation for the nomisweb api can be found at: 
https://www.nomisweb.co.uk/api/v01/help

Examples of how to use the jsonlite package can be found at:
https://cran.r-project.org/web/packages/jsonlite/vignettes/json-apis.html

Links for information on SDMX: 
https://sdmx.org
https://sdmx.org/wp-content/uploads/SDMX_2-1_User_Guide_draft_0-1.pdf
https://sdmx.org/?page_id=4500 - tools for working with sdmx


## TODO

### Investigate using jsonstat 
Nomisweb offers data downloads in multiple formats:

* HTML
* XLSX
* csv 
* xml
* JSON (a proprietary flavour)
* JSONstat (a standard https://json-stat.org/)

csv is perhaps the most obvious and easy to understand way to extract data and as such the scripts will target this first. However JSON stat is used by multiple other statistics bodies. As such it might be interesting to understand how it works. 

There is an R library for parsing jsonstat:
```{r load_libs_2, echo=FALSE}
library(rjstat)
```


# Use 1 - Get List of Available Datasets
Step 1 - Get a list of the available datasets, obtaining
* ID for use with the API
* A Description
* A Name

This api request gets you a 'large list' which contains information about all the datasets on nomisweb as a JSON.

To understand what this contains you can look at the HTML version at: https://www.nomisweb.co.uk/api/v01/dataset/def.htm

* Structure
** "header"         
** "keyfamilies"    
** "xmlns"          
** "common"         
** "structure"      
** "xsi"            
** "schemalocation"

```{r}
nomisDatasets <- fromJSON("https://www.nomisweb.co.uk/api/v01/dataset/def.sdmx.json")

```

```{r}
datasets<-nomisDatasets$structure$keyfamilies$keyfamily %>% select(id,description,name)

rm(nomisDatasets)
```


A call to https://www.nomisweb.co.uk/api/v01/dataset/<datasetid>/geography.def.sdmx.json will render a list of top level areas for which the dataset is available.

This will hopefully include: 
* 2092957697	United Kingdom

A further call to https://www.nomisweb.co.uk/api/v01/dataset/<datasetid>/geography/2092957697.def.json will then render a list of the geography types which are available within the UK top level area.

We are interested in TYPE464 - local authorities: district / unitary (prior to April 2015) within United Kingdom.

So, what we want to do is iterate over the data frame, adding a new columns which check:
* 1 - if the dataset is available for the UK 
* 2 - if the dataset is available in Type464 output areas

```{r}
# Define a function which queries the nomisweb api for a given dataset - asking 

# What top level areas are available?	/api/v01/dataset/NM_1_1/geography.def.sdmx.json

checkuk <- function(dset){
  
  countrieslist <- fromJSON(
      paste0("https://www.nomisweb.co.uk/api/v01/dataset/",
             dset,
             "/geography.def.sdmx.json"))$structure$codelists$codelist$code[[1]]
  
  # boolean check of the first item in the top levels region list against the UK region code
  UK<-(countrieslist$value[1]==2092957697)
  print(paste(dset, UK, sep=' '))
  return(UK)    
}

```

Check which datasets are in the UK
```{r}
dsetinUK<-lapply(datasets$id, checkuk)
```

Once we have this vector telling us if it is in the UK or not we can add this to our datasets dataframe
```{r}
dsetinUK<-data_frame(dsetinUK)
datasets2 <- cbind(datasets, dsetinUK)
```


Now that we know which datasets are available for the UK we want to see which Area types are available for each UK dataset.
```{r}

getGeogTable <- function(dset){
  
  Arealist <- fromJSON(
      paste0("https://www.nomisweb.co.uk/api/v01/dataset/",
             dset,
             "/geography/2092957697.def.sdmx.json"))
      
  return(Arealist)    
}

```

We can then use our getGeogTable function to get a table of the geographies available for a particular variable 
```{r}
geogTable<-getGeogTable("NM_1_1")$structure$codelists$codelist$code[[1]]

# Check if Type 464 - Pre 2015 Local Areas is in our geogTable
if("2092957697TYPE464" %in% unlist(geogTable$value)){
  # If it is, we want to download the dataset
  print("yeah")
}
```

# Use 2 - Download a Dataset
If the dataset is available for the UK at type464 level then we probably want to download it.
```{r}
getNomisData <- function(dset){
  
  
}

```

