---
title: "SDMX"
author: "Matthew Malcher"
date: "17/11/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#  Load Libraries
There is an R library for working with SDMX. NOMIS is included as a predefined data source! This looks like it might make things a bit easier, but I cant figure out how to take advantage of it. Fortunately you can also buld your own URL's and feed them to the `readSDMX` command provided by the library.
```{r }
library(rsdmx)
providers <- getSDMXServiceProviders();
as.data.frame(providers) %>% filter(agencyId=='NOMIS')
```

At this point I will also load a few other useful libraries:
```{r}
library(dplyr)
library(tidyr)
```


There are some common bits of info used 
```{r}
# Define API access point
ap<-'https://www.nomisweb.co.uk/api/v01/dataset'
qformat <-'def.sdmx.xml'

```



Now, lets get started with the nomisweb API!

With the following commands we can get a data frame with details of all the datasets available on NOMIS
```{r}
queryUrl <- paste(ap,qformat,sep='/')
queryUrl
NomisDatasets <- as.data.frame(readSDMX(queryUrl))
```


For a particular data item, see further info
```{r}
dataset<-'NM_1_1'
queryUrl<-paste(ap,dataset,qformat,sep='/')
SingleDataset <- as.data.frame(readSDMX(queryUrl))
SingleDataset
```



Get a Table of available regions for a dataset:
```{r}
dataset<-'NM_1_1'
qtype<-'geography.'
queryUrl<-paste0(paste(ap,dataset,qtype,sep='/'),qformat)
queryUrl
AvailableRegions <- as.data.frame(readSDMX(queryUrl))
AvailableRegions
```

Get a Table of available output areas for a dataset
```{r}
dataset<-'NM_1_1'
qtype<-'geography'
geography<-'2092957697.'
queryUrl<-paste0(paste(ap,dataset,qtype,geography,sep='/'),qformat)
queryUrl
AvailableGeographies <- as.data.frame(readSDMX(queryUrl))
AvailableGeographies
```

https://www.nomisweb.co.uk/forum/posts.aspx?tID=555&fID=2

Get List of output areas for a given geography type
```{r}
dataset<-'NM_1_1'
qtype<-'geography'
geography<-'2092957697'
OAtype<-'TYPE464'

queryUrl<-paste0(paste(ap,dataset,qtype,geography,OAtype,sep='/'),qformat)

OutputAreas <- as.data.frame(readSDMX(queryUrl))
OutputAreas

```

Get a Table of available sex information for a dataset
```{r}

dataset<-'NM_1_1'
qtype<-'sex'
queryUrl<-paste0(paste(ap,dataset,qtype,sep='/'),qformat)
queryUrl
AvailableSexes <- as.data.frame(readSDMX(queryUrl))
AvailableSexes
```

Get the measures available for a given geography - these are denominators - i.e. out of the number of claimints, the number of people living there etc.
```{r}
dataset<-'NM_1_1'
qtype<-'measures'
geography<-'2092957697'
OAtype<-'TYPE464'
queryUrl<-paste0(paste(ap,dataset,qtype,geography,OAtype,sep='/'),qformat)
queryUrl
Measures <- as.data.frame(readSDMX(queryUrl))
Measures
```

Get available 'items' for a specified dataset (i.e. the variables)
```{r}
dataset<-'NM_1_1'
qtype<-'item'
queryUrl<-paste0(paste(ap,dataset,qtype,sep='/'),qformat)
queryUrl
AvailableItems <- as.data.frame(readSDMX(queryUrl))
AvailableItems

```

/api/v01/dataset/NM_1_1.generic.sdmx.xml?geography=2038432081&sex=5&item=1&measures=20100&time=latest

Get available 'items' for a specified dataset (i.e. the variables)
```{r}

dataset<-'NM_1_1'

geography='2092957697TYPE464'
sex='7'
item='1'
measures='20100'
time='latest'

query<-paste0('?geography=',geography,'&sex=',sex,'&item=',item,'&measures=',measures,'&time=latest')

#Dont need to have all items  
query<-paste0('?geography=',geography,'&item=',item,'&measures=',measures,'&time=latest')

format<-'.generic.sdmx.xml'

file<-paste0(paste(ap,dataset,sep='/'),format,query)

sdmx <- readSDMX(file)

df <- as.data.frame(sdmx)

df
```

```{r}

df %>%
  filter(SEX=='7') %>%
  join() %>%
  
```

